\chapter{The rigidity lemma and its Milne \S I.1--I.3 corollaries}
\label{chap:AbelianVarietyRigidity}
\label{chap:avr_for_rr}
% archon:covers AlgebraicJacobian/RigidityLemma.lean
% Local macro: diagrammatic (Lean `≫`) composition. Defined locally because this
% chapter is the writer's only editable file; flagged in the writer report for promotion
% to macros/common.tex if desired. \providecommand is a no-op if it is later added globally.
\providecommand{\fatsemi}{\mathbin{\,;\,}}
% STRATEGY NOTE. This chapter is the dedicated home of the GENERAL Mumford/Milne \S I.1--I.3
% rigidity-lemma infrastructure. It is 1:1 with the upstream Lean file
% AlgebraicJacobian/AbelianVarietyRigidity.lean (the RigidityLemma.lean material), which imports
% AlgebraicJacobian.Genus and is imported by AlgebraicJacobian.Jacobian. The material is consumed
% by the positive-genus Albanese universal property of Pic^0_{C/k}: the rigidity lemma and its
% corollaries (a pointed regular map of abelian varieties is a homomorphism; additivity over a
% product) are exactly the inputs the Albanese factorisation needs.
This chapter develops the general \emph{Rigidity Lemma} of Mumford and its Milne \S I.1--I.3
corollaries. These are the elementary, cohomology-free building blocks of abelian-variety
theory: the lemma collapses a morphism out of a product when one fibre is constant, and from it
follow the basic structural facts that a pointed regular map of abelian varieties is a group
homomorphism and that \(\mathrm{Hom}(-, A)\) is additive over a product. This infrastructure is
consumed by the positive-genus Albanese universal property of \(\Pic^0_{C/k}\): the uniqueness
and homomorphism content of the Albanese factorisation rest on exactly these rigidity
corollaries. (When \(\genus C = 0\) the Jacobian \(\Pic^0_{C/k}\) is automatically \(\Spec k\),
since \(H^1(C, \mathcal O_C) = 0\), so the genus-\(0\) case is the degenerate instance of the
same universal property and needs no separate treatment.)
\medskip
The \emph{Rigidity Lemma} (\cref{thm:rigidity_lemma}, Mumford, Form~I; equivalently Milne's
Rigidity Theorem~1.1) is an elementary properness/closed-map argument --- it uses neither the
theorem of the cube nor any cohomology, is the entry point for formalisation, and is now proven
axiom-clean. From the Rigidity Lemma, Milne's \S I.1--I.3 derive the structural corollaries
\emph{without} the theorem of the cube. The decisive input is that a complete first factor makes
the second projection \(p_2 \colon X \times Y \to Y\) a closed map; combined with the universal
property of the product this forces a morphism whose restriction to one fibre is constant to
factor through that projection.
The chapter then records the two corollaries used by the Albanese construction. Additivity of
\(\mathrm{Hom}(-, A)\) over a product (\cref{lem:hom_additivity_over_product}, Milne
Corollary~1.5) --- whose Lean signature requires only the \emph{first} factor complete --- is
obtained directly from the lemma applied to the defect map \((x, y) \mapsto f(x, y) - f(x, y_0)
- f(x_0, y) + f(x_0, y_0)\), which is constant on each axis. From it follows Milne
Corollary~1.2, that a pointed regular map of abelian varieties is a homomorphism
(\cref{lem:av_regular_map_is_hom}); this is the corollary the positive-genus Albanese
universal property consumes. The theorem of the cube (Milne \S I.5) is \emph{not} used anywhere
on this path (see \cref{rmk:cube_not_needed}); the chain invokes neither Serre duality nor the
FGA \(\Pic\) representability engine. Milne's Theorem~3.2 (rational-map extension,
\cref{thm:rational_map_to_av_extends}) is a Route-A-only item (the Albanese universal property),
recorded here for completeness (\cref{rmk:thm32_codim1_mathlib_gap}).
\medskip
Throughout, \(\bar k\) is an algebraically closed field of arbitrary characteristic. An
\emph{abelian variety} over \(\bar k\) is a complete (proper) irreducible group variety; in the
project's Lean encoding \(A\) is an object of \(\mathrm{Over}\,(\Spec \bar k)\) carrying
\(\mathtt{GrpObj}\), \(\mathtt{IsProper}\), \(\mathtt{Smooth}\), and \(\mathtt{GeometricallyIrreducible}\).
\section{The Rigidity Lemma}
\begin{lemma}
\leanok
[Rigidity Lemma (Mumford, Form I)]
  \label{thm:rigidity_lemma}
  \uses{lem:rigidity_eqOn_dense_open, lem:rigidity_core, lem:rigidity_snd_lift}
  
  \lean{AlgebraicGeometry.rigidity_lemma}
  % SOURCE: [Mumford, Abelian Varieties], Rigidity Lemma (Form I.), Ch.~II \S4, book p.~43 (read from references/mumford-abelian-varieties.pdf, PDF page 54)
  % SOURCE QUOTE: "RIGIDITY LEMMA. (Form I.) Let \(X\) be a complete variety, \(Y\) and \(Z\)
  % any varieties, and \(f \colon X \times Y \to Z\) a morphism such that for some
  % \(y_0 \in Y\), \(f(X \times \{y_0\})\) is a single point \(z_0\) of \(Z\). Then there is a
  % morphism \(g \colon Y \to Z\) such that if \(p_2 \colon X \times Y \to Y\) is the
  % projection, \(f = g \circ p_2\)."
  \textit{Source: Mumford, Abelian Varieties, Ch.~II \S4, Rigidity Lemma (Form~I), p.~43.}
  Let \(X\) be a complete variety, and let \(Y\), \(Z\) be any varieties over \(\bar k\). Let
  \(f \colon X \times Y \to Z\) be a morphism such that for some \(y_0 \in Y\) the image
  \(f(X \times \{y_0\})\) is a single point \(z_0 \in Z\). Then there is a morphism
  \(g \colon Y \to Z\) with \(f = g \circ p_2\), where \(p_2 \colon X \times Y \to Y\) is the
  projection; that is, \(f\) depends only on the \(Y\)-coordinate, and \(g(y) = f(x_0, y)\) for any
  fixed \(x_0 \in X\).
\end{lemma}
% SOURCE QUOTE PROOF: "PROOF. Choose any point \(x_0 \in X\), and define \(g \colon Y \to Z\) by
% \(g(y) = f(x_0, y)\). Since \(X \times Y\) is a variety, to show that \(f = g \circ p_2\), it is
% sufficient to show that these morphisms coincide on some open subset of \(X \times Y\). Let
% \(U\) be an affine open neighbourhood of \(z_0\) in \(Z\), \(F = Z - U\), and \(G = p_2(f^{-1}(F))\);
% then \(G\) is closed in \(Y\) since \(X\) is complete and hence \(p_2\) is a closed map. Further
% \(y_0 \notin G\) since \(f(X \times \{y_0\}) = \{z_0\}\). Therefore \(Y - G = V\) is a non-empty
% open subset of \(Y\). For each \(y \in V\), the complete variety \(X \times \{y\}\) gets mapped by
% \(f\) into the affine variety \(U\), and hence to a single point of \(U\). But this means that for
% any \(x \in X\), \(y \in V\), \(f(x, y) = f(x_0, y) = g \circ p_2(x, y)\), and this proves our
% assertion."
\begin{proof}
\leanok
% NOTE (iter-159 review): the iter-157 unsoundness REMAINS REPAIRED. The collapse hypothesis
% (_hf : f(X x {y0}) = {z0}) is threaded through all three helpers (rigidity_lemma ->
% rigidity_core -> rigidity_eqOn_dense_open) and GENUINELY CONSUMED in the body of
% rigidity_eqOn_dense_open (proves y0 \notin G, the load-bearing non-emptiness of V := Y - G).
% The counterexample f = fst is excluded (under f = fst, _hf forces X to collapse to a point).
% rigidity_eqOn_dense_open is sorry-free in its OWN body (hfib, the pullback-fibre fact over the
% k-bar-point y0, is closed). rigidity_snd_lift and snd_left_isClosedMap are axiom-clean (no sorryAx).
% NOTE (iter-160): the \uses edges run FORWARD along the true Lean dependency order
% (rigidity_lemma -> rigidity_eqOn_dense_open -> rigidity_eqOn_saturated_open_to_affine), so the
% not-proven status of the residual sorry propagates up and de-launders BOTH thm:rigidity_lemma and
% lem:rigidity_eqOn_dense_open. No backward edge from the leaf lemmas back up the chain.
% NOTE (iter-161): bridge 2 (rigidity_eqOn_saturated_open_to_affine) was DECOMPOSED into two named
% top-level sub-lemmas along the route-B (cohomology-free) split:
%   - Step 2, morphism_eq_of_eqAt_closedPoints: "two morphisms agreeing at every closed point of a
%     reduced Jacobson source into a separated target are equal" -- PROVEN axiom-clean (the
%     dense-closed-points coproduct probe + ext_of_isDominant; the connective Mathlib lacks).
%   - Step 1, rigidity_eqAt_closedPoint_of_proper_into_affine: Mumford's per-closed-slice constancy
%     -- the genuinely-deep residual, still `sorry`. THIS is now the chain's lone deep residual
%     (bridge 2's body is real assembly: Step 2 proven, the JacobsonSpace U instance now derivable).
% SIGNATURE CHANGE (iter-161, plan-authorized): Step 2's globalisation needs the closed points of
% the saturated open U to be DENSE, i.e. U Jacobson, which over Spec k-bar follows from (X tensor Y)
% being LOCALLY OF FINITE TYPE. So [LocallyOfFiniteType (X tensor Y).hom] is now carried on FIVE
% chain lemmas (rigidity_lemma, rigidity_core, rigidity_eqOn_dense_open,
% rigidity_eqOn_saturated_open_to_affine, rigidity_eqAt_closedPoint_of_proper_into_affine), in
% addition to [IsAlgClosed kbar]. Free downstream: curves and abelian varieties are of finite type
% over k-bar. The earlier "[IsAlgClosed] is the only added instance" claim is hereby retired.
  Fix any point \(x_0 \in X\) and define \(g \colon Y \to Z\) by \(g(y) = f(x_0, y)\) (the
  composite of \(y \mapsto (x_0, y)\) with \(f\)). Since \(X \times Y\) is a variety (in
  particular reduced and irreducible, so separated and with no embedded structure), to prove
  the equality of morphisms \(f = g \circ p_2\) it suffices to prove they agree on a non-empty
  open subset of \(X \times Y\).
  Let \(U\) be an affine open neighbourhood of \(z_0\) in \(Z\), put \(F = Z - U\) (a closed set),
  and set \(G = p_2(f^{-1}(F))\). Because \(X\) is complete, the projection
  \(p_2 \colon X \times Y \to Y\) is a \emph{closed} map (this is the defining property of
  completeness/properness); hence \(G\), the image of the closed set \(f^{-1}(F)\), is closed in
  \(Y\). The hypothesis \(f(X \times \{y_0\}) = \{z_0\} \subseteq U\) gives
  \(f^{-1}(F) \cap (X \times \{y_0\}) = \varnothing\), so \(y_0 \notin G\). Therefore
  \(V := Y - G\) is a non-empty open subset of \(Y\).
  For each \(y \in V\) we have \(X \times \{y\} \cap f^{-1}(F) = \varnothing\), i.e.
  \(f(X \times \{y\}) \subseteq U\). Thus the complete (proper) variety \(X \times \{y\}\) is
  mapped by \(f\) into the \emph{affine} variety \(U\). A proper connected variety mapping to an
  affine variety has image a single point (a global regular function on a complete connected
  variety is constant). Hence \(f(X \times \{y\})\) is one point, necessarily
  \(f(x_0, y) = g(y)\). So for all \(x \in X\) and \(y \in V\),
  \[ f(x, y) = f(x_0, y) = g(y) = (g \circ p_2)(x, y). \]
  The two morphisms \(f\) and \(g \circ p_2\) thus agree on the non-empty open
  \(X \times V\), which is dense in the irreducible \(X \times Y\); as \(Z\) is separated they
  agree everywhere. \qedhere
\end{proof}
\begin{remark}
  \label{rmk:rigidity_lemma_cube_free}
  The proof of \cref{thm:rigidity_lemma} uses \emph{only} (i) that completeness of \(X\) makes
  \(p_2\) a closed map, and (ii) that a proper connected variety has no nonconstant map to an
  affine variety. It uses neither the theorem of the cube nor any cohomology, and is valid in
  arbitrary characteristic. This is the designated formalisation entry point: it is the
  lowest-prerequisite link of the chain. (Mumford gives this as the geometric heart of the
  ``second proof'' of commutativity of an abelian variety; Milne packages the same statement
  as the Rigidity Theorem~1.1, requiring \(V\) complete and \(V \times W\) geometrically
  irreducible.)
\end{remark}
\begin{remark}
  \label{rmk:rigidity_lemma_decomposition}
  The Lean formalisation of \cref{thm:rigidity_lemma} splits into three helper steps, mirroring
  the proof above: \texttt{rigidity\_snd\_lift} (the cartesian-monoidal algebra identifying the
  collapsed map \((x, y) \mapsto (x_0, y)\) followed by \(f\) with \(g \circ p_2\));
  \texttt{rigidity\_core} (the scheme-level gluing: two maps out of the geometrically irreducible
  reduced \(X \times Y\) into the separated \(Z\) that agree on a non-empty open subset agree
  everywhere, by separatedness); and \texttt{rigidity\_eqOn\_dense\_open}
  (\cref{lem:rigidity_eqOn_dense_open}, the genuine geometry: the construction of the non-empty
  open \(X \times V\) and the slice-constancy of \(f\) on it). The collapse hypothesis
  \(f(X \times \{y_0\}) = \{z_0\}\) must be threaded through \emph{all three} helpers, since it is
  what makes \(V\) non-empty; a decomposition that drops it from the lower helpers is unsound
  (see \cref{lem:rigidity_eqOn_dense_open}).
  \emph{Status (iter-161).} Bridge~1 (the closed-map step) is \textbf{built} as the axiom-clean
  helper \texttt{snd\_left\_isClosedMap}. The fibre-over-\(y_0\) fact is \textbf{closed} (via the
  coarse \texttt{PullbackCarrier} layer), so \texttt{rigidity\_eqOn\_dense\_open} is
  \texttt{sorry}-free in its own body. Bridge~2 (slice-constancy on the saturated open \(U\),
  \cref{lem:rigidity_eqOn_saturated_open_to_affine}) has now been \textbf{decomposed} along its
  cohomology-free route-(c) split into two named sub-lemmas: the globalisation
  \cref{lem:morphism_eq_of_eqAt_closedPoints} (Step~2, ``agree at every closed point of a reduced
  Jacobson source \(\Rightarrow\) equal'') is \textbf{proven axiom-clean}, and the per-closed-slice
  constancy \cref{lem:rigidity_eqAt_closedPoint_of_proper_into_affine} (Step~1, Mumford's slice
  argument) is the chain's \emph{single genuinely-deep residual} \texttt{sorry}. The body of
  \cref{lem:rigidity_eqOn_saturated_open_to_affine} is therefore real assembly (Step~2 over Step~1).
  % NOTE: (iter-162 review) STALE as of iter-162 — Step~1 is now PROVEN and the WHOLE
  % Rigidity-Lemma chain (thm:rigidity_lemma … lem:rigidity_eqAt_closedPoint_of_proper_into_affine,
  % via the new helper lem:isIntegral_of_retract_of_integral) is sorry-free + axiom-clean
  % ({propext, Classical.choice, Quot.sound}, re-verified). The "single genuinely-deep residual
  % sorry" wording here and in the proof blocks of lem:rigidity_eqOn_dense_open / …saturated… /
  % …eqAt_closedPoint… should be refreshed by the plan/blueprint-writer to "chain closed iter-162".
  This route requires the base field to be algebraically closed \emph{and} the source to be of finite
  type: Step~2 needs the closed points of \(U\) to be dense (the Jacobson-space property), which over
  the algebraically closed \(\bar k\) holds because \(X \times Y\) is locally of finite type.
  Accordingly the formalization carries \textbf{both} \([\mathtt{IsAlgClosed}\ \bar k]\) and
  \([\mathtt{LocallyOfFiniteType}\ (X \otimes Y).\mathrm{hom}]\) across the chain lemmas
  (\texttt{rigidity\_lemma}, \texttt{rigidity\_core}, \texttt{rigidity\_eqOn\_dense\_open},
  \texttt{rigidity\_eqOn\_saturated\_open\_to\_affine},
  \texttt{rigidity\_eqAt\_closedPoint\_of\_proper\_into\_affine}); the earlier claim that
  \([\mathtt{IsAlgClosed}\ \bar k]\) is the \emph{only} added instance is retired. Both cost nothing
  downstream: curves and abelian varieties are of finite type over the algebraically closed \(\bar k\),
  and the consumers already assume it. The relative Stein-factorisation / proper-pushforward
  \(f_*\mathcal O = \mathcal O\) framing is a genuine Mathlib cohomology gap and is \emph{deliberately
  avoided}.
\end{remark}
\begin{lemma}\leanok
[Cartesian-monoidal collapse identity]
  \label{lem:rigidity_snd_lift}
  \lean{AlgebraicGeometry.rigidity_snd_lift}
  Let \(X\), \(Y\) be objects over \(\bar k\) and fix a \(\bar k\)-point \(x_0\) of \(X\).
  Post-composing the second projection \(p_2 \colon X \times Y \to Y\) with the slice section
  \(y \mapsto (x_0, y)\) equals the ``collapse the \(X\)-coordinate onto \(x_0\)'' endomorphism
  \((x, y) \mapsto (x_0, y)\) of \(X \times Y\). This is pure cartesian-monoidal algebra (no
  geometry): the product universal property distributes \(p_2\), the \(Y\)-component simplifies by
  the identity law, and the \(X\)-component collapses by uniqueness of maps to the terminal object.
\end{lemma}
\begin{lemma}\leanok
[Bridge~1: completeness makes the projection a closed map]
  \label{lem:snd_left_isClosedMap}
  \lean{AlgebraicGeometry.snd_left_isClosedMap}
  Let \(X\) be complete (proper) over \(\bar k\) and \(Y\) any object over \(\bar k\). Then the
  underlying topological map of the second projection \(p_2 \colon X \times Y \to Y\) is a
  \emph{closed} map. This is Mumford's ``completeness of \(X\) makes \(p_2\) a closed map'': the
  projection is the base change of \(X \to \Spec \bar k\) along \(Y \to \Spec \bar k\), properness
  of \(X\) gives universal closedness, and universal closedness is stable under base change, so the
  base-changed projection is universally closed and hence a closed map. Characteristic-free; no
  theorem of the cube, no cohomology.
\end{lemma}
\begin{lemma}\leanok
[Scheme-level gluing core of the Rigidity Lemma]
  \label{lem:rigidity_core}
  \lean{AlgebraicGeometry.rigidity_core}
  \uses{lem:rigidity_eqOn_dense_open}
  Let \(\bar k\) be algebraically closed, \(X\) proper, \(X \times Y\) geometrically irreducible,
  reduced and locally of finite type over \(\bar k\), and \(Z\) separated. Let
  \(f \colon X \times Y \to Z\) collapse the slice \(X \times \{y_0\}\) to a single point \(z_0\).
  Then \(f\) equals its composite with the collapse endomorphism \((x, y) \mapsto (x_0, y)\), i.e.\
  \(f = \mathtt{retract} \fatsemi f\). The proof takes the non-empty open \(U\) on which \(f\) and
  \(\mathtt{retract} \fatsemi f\) agree (\cref{lem:rigidity_eqOn_dense_open}); since \(X \times Y\)
  is geometrically irreducible over the one-point base \(\Spec \bar k\), it is irreducible, so the
  non-empty open \(U\) is dense and its inclusion is dominant; the dominant-source/separated-target
  rigidity handle then promotes agreement on \(U\) to equality of the two morphisms everywhere.
\end{lemma}
\begin{lemma}
\leanok
[Dense-open agreement: the geometric heart of the Rigidity Lemma]
  \label{lem:rigidity_eqOn_dense_open}
  \lean{AlgebraicGeometry.rigidity_eqOn_dense_open}
  \uses{lem:rigidity_eqOn_saturated_open_to_affine, lem:snd_left_isClosedMap}
  % SOURCE: [Mumford, Abelian Varieties], Rigidity Lemma (Form I.), Ch.~II \S4, book p.~43 (read from references/mumford-abelian-varieties.pdf, PDF page 54)
  % SOURCE QUOTE: "RIGIDITY LEMMA. (Form I.) Let \(X\) be a complete variety, \(Y\) and \(Z\)
  % any varieties, and \(f \colon X \times Y \to Z\) a morphism such that for some
  % \(y_0 \in Y\), \(f(X \times \{y_0\})\) is a single point \(z_0\) of \(Z\). Then there is a
  % morphism \(g \colon Y \to Z\) such that if \(p_2 \colon X \times Y \to Y\) is the
  % projection, \(f = g \circ p_2\)."
  \textit{Source: Mumford, Abelian Varieties, Ch.~II \S4, Rigidity Lemma (Form~I), p.~43 (the
  dense-open construction is the body of that proof).}
  Let \(X\) be a complete variety and \(Y\), \(Z\) any varieties over \(\bar k\). Let
  \(f \colon X \times Y \to Z\) be a morphism, fix a point \(x_0 \in X\), and suppose given
  \(y_0 \in Y\), \(z_0 \in Z\) with the \emph{collapse hypothesis}
  \[ f(X \times \{y_0\}) = \{z_0\} \qquad
     \text{(in Lean: } \mathtt{lift}\,(\mathbf 1_X)\,(\mathtt{toUnit}\,X \fatsemi y_0) \fatsemi f
     = \mathtt{toUnit}\,X \fatsemi z_0\text{).} \]
  Then there is a non-empty open subset \(U \subseteq X \times Y\) on which \(f\) agrees, as a
  scheme morphism, with the collapsed map \(\mathtt{retract} \fatsemi f\), where
  \(\mathtt{retract} := \mathtt{lift}\,(\mathtt{toUnit}\,(X \times Y) \fatsemi x_0)\,
  (p_2 \colon X \times Y \to Y)\) is the ``\((x, y) \mapsto (x_0, y)\)'' endomorphism. Concretely
  \(U = X \times V\) with \(V := Y - G\), \(G = p_2(f^{-1}(Z - U_0))\) for an affine open
  neighbourhood \(U_0\) of \(z_0\), and on \(U\) one has \(f(x, y) = f(x_0, y)\).
  \medskip
  \noindent\textbf{The collapse hypothesis is load-bearing.} The non-emptiness of \(V\) is exactly
  the assertion \(y_0 \notin G\), and this holds \emph{because} \(f(X \times \{y_0\}) = \{z_0\}
  \subseteq U_0\) keeps the slice over \(y_0\) inside the affine \(U_0\), away from \(F = Z - U_0\). A
  version of this lemma \emph{without} the collapse hypothesis is false: for \(X = Y = Z\) and
  \(f = p_1\) the first projection, \(f\) and \(\mathtt{retract} \fatsemi f \colon (x, y) \mapsto x_0\)
  agree on no non-empty open (the agreement locus is contained in \(\{x = x_0\}\), which has empty
  interior in the irreducible \(X \times Y\)). Therefore the formal target \emph{must} carry the
  collapse hypothesis as an explicit antecedent; dropping it (as an earlier decomposition
  silently did) makes the deferred goal unsatisfiable.
  \medskip
  \noindent\textbf{Formalization note: algebraically closed base, and locally-of-finite-type
  source.} The prose already takes \(\bar k\) algebraically closed, but the formal statement must
  \emph{encode} two properties. First, the cohomology-free proof of the slice-constancy bridge
  (below) needs each per-closed-point residue field \(\kappa(y)\) to be \(\bar k\) itself, so that a
  proper integral slice has global sections \(\Gamma = \bar k\) and maps to a single \(\bar k\)-point of
  the affine \(U_0\); this requires \([\mathtt{IsAlgClosed}\ \bar k]\). Second, the globalisation of the
  per-slice agreement (Step~2 of the bridge, \cref{lem:morphism_eq_of_eqAt_closedPoints}) needs the
  closed points of the saturated open \(U\) to be \emph{dense} --- i.e.\ \(U\) to be a Jacobson space.
  Over the algebraically closed \(\bar k\) this holds precisely because \(X \times Y\) is \emph{locally of
  finite type} over \(\bar k\) (\texttt{LocallyOfFiniteType.jacobsonSpace} transports the Jacobson
  property of \(\Spec \bar k\), and \texttt{closure\_closedPoints} then gives density); this requires
  \([\mathtt{LocallyOfFiniteType}\ (X \otimes Y).\mathrm{hom}]\). Accordingly the formalization carries
  \emph{both} \([\mathtt{IsAlgClosed}\ \bar k]\) and
  \([\mathtt{LocallyOfFiniteType}\ (X \otimes Y).\mathrm{hom}]\) on this lemma (in addition to
  \([\mathtt{Field}\ \bar k]\)), and propagates the same two instances to \cref{thm:rigidity_lemma},
  its core helper \texttt{rigidity\_core}, and the bridge sub-lemmas
  \cref{lem:rigidity_eqOn_saturated_open_to_affine} and
  \cref{lem:rigidity_eqAt_closedPoint_of_proper_into_affine}. Both cost nothing downstream: the
  abelian-variety consumers of the rigidity corollaries work over an algebraically closed field
  and \emph{already} assume \([\mathtt{IsAlgClosed}\ \bar k]\) --- the base-field variable is
  literally named \(\mathtt{kbar}\) --- and curves and abelian varieties are of finite type over
  \(\bar k\), so the locally-of-finite-type instance is automatic there.
\end{lemma}
% SOURCE QUOTE PROOF: "PROOF. Choose any point \(x_0 \in X\), and define \(g \colon Y \to Z\) by
% \(g(y) = f(x_0, y)\). Since \(X \times Y\) is a variety, to show that \(f = g \circ p_2\), it is
% sufficient to show that these morphisms coincide on some open subset of \(X \times Y\). Let
% \(U\) be an affine open neighbourhood of \(z_0\) in \(Z\), \(F = Z - U\), and \(G = p_2(f^{-1}(F))\);
% then \(G\) is closed in \(Y\) since \(X\) is complete and hence \(p_2\) is a closed map. Further
% \(y_0 \notin G\) since \(f(X \times \{y_0\}) = \{z_0\}\). Therefore \(Y - G = V\) is a non-empty
% open subset of \(Y\). For each \(y \in V\), the complete variety \(X \times \{y\}\) gets mapped by
% \(f\) into the affine variety \(U\), and hence to a single point of \(U\). But this means that for
% any \(x \in X\), \(y \in V\), \(f(x, y) = f(x_0, y) = g \circ p_2(x, y)\), and this proves our
% assertion."
\begin{proof}
  \leanok
  Let \(U_0\) be an affine open neighbourhood of \(z_0\) in \(Z\), put \(F = Z - U_0\), and set
  \(G = p_2(f^{-1}(F))\). Because \(X\) is complete, the projection \(p_2 \colon X \times Y \to Y\) is a
  \emph{closed} map, so \(G\) (the image of the closed set \(f^{-1}(F)\)) is closed in \(Y\). The
  collapse hypothesis \(f(X \times \{y_0\}) = \{z_0\} \subseteq U_0\) gives
  \(f^{-1}(F) \cap (X \times \{y_0\}) = \varnothing\), hence \(y_0 \notin G\); this is the only place
  the collapse hypothesis is used, and it is essential, for \(V := Y - G\) is non-empty precisely
  because \(y_0 \in V\). For each \(y \in V\) the proper variety \(X \times \{y\}\) is mapped by \(f\)
  into the affine \(U_0\), hence to a single point, necessarily \(f(x_0, y)\). So for all \(x \in X\),
  \(y \in V\), we have \(f(x, y) = f(x_0, y) = (\mathtt{retract} \fatsemi f)(x, y)\), i.e.\ \(f\) and
  \(\mathtt{retract} \fatsemi f\) agree on the non-empty open \(U := X \times V\). (In the Lean
  encoding the two slice-constancy bridges are the closed-map step from properness and the
  ``proper-into-affine has single-point image'' step; see the docstring of
  \texttt{rigidity\_core}.)
  \medskip
  \noindent\textbf{Formalization notes (the concrete char-free Mathlib route).} The textbook
  proof above is realised in Lean by three pieces. They cite Mathlib lemma names as the chosen
  route; none uses cohomology, and the relative Stein-factorisation framing is deliberately
  \emph{avoided} (see below).
  \emph{Bridge~1 (the closed-map step) --- BUILT.} ``\(X\) complete \(\Rightarrow p_2\) closed'' is
  the axiom-clean helper \texttt{snd\_left\_isClosedMap} (iter-158). It is obtained from
  \texttt{IsProper.toUniversallyClosed} on \(X\), base-changed across the canonical pullback square
  to make \(p_2\) universally closed, using that the monoidal projection \emph{is} the chosen
  pullback projection: \texttt{Over.snd\_left} rewrites \((\mathtt{snd}\,X\,Y).\mathtt{left}\) to
  \(\mathtt{Limits.pullback.snd}\ X.\mathtt{hom}\ Y.\mathtt{hom}\) exactly. This step is already
  proven.
  \emph{The fibre fact (non-emptiness of \(V\)) --- RESOLVED route, char-free.} That the slice
  \(X \times \{y_0\}\) over the \(\bar k\)-rational point \(y_0\) is exactly the image of the slice
  section \(s \colon X \to X \times Y\) is formalised \emph{without} the residue-field /
  \texttt{Triplet} / tensor-product machinery (that fine \texttt{carrierEquiv} layer is the wrong
  granularity and is to be avoided). One pastes the identity square for the section against the
  canonical pullback square (\texttt{IsPullback.of\_right},
  \texttt{IsPullback.of\_horiz\_isIso}; the section identities \(s \fatsemi p_1 = \mathbf 1\) and
  \(s \fatsemi p_2 = X.\mathtt{hom} \fatsemi y_0\) come from \texttt{lift\_fst}/\texttt{lift\_snd}
  with \texttt{Over.w} on the unit) and reads off the fibre with the \emph{coarse} topological
  layer of \texttt{PullbackCarrier}, namely
  \texttt{AlgebraicGeometry.Scheme.Pullback.image\_preimage\_eq\_of\_isPullback}. This needs no
  \([\mathtt{IsAlgClosed}]\).
  \emph{Bridge~2 (slice-constancy / the agreement equation) --- RESOLVED route, cohomology-FREE.}
  ``Proper connected slice into affine \(\Rightarrow\) single point, glued to a morphism equality on
  \(U\)'' is realised by a \textbf{global-sections + per-closed-point} route, \emph{not} by any
  relative Stein-factorisation or proper-pushforward \(f_*\mathcal O = \mathcal O\) statement: the
  latter is a genuine Mathlib gap (no Stein factorisation, no \(H^0\) base change), and route~(c)
  exists precisely to avoid coherent cohomology, so it must \emph{not} be attempted. The concrete
  stack is: for each closed point \(y \in V\) one has \(\kappa(y) = \bar k\) (alg-closed, finite type);
  the slice \(X_y\) is proper integral, so its global sections form a field
  (\texttt{isField\_of\_universallyClosed}) module-finite over \(\bar k\)
  (\texttt{finite\_appTop\_of\_universallyClosed}), and algebraic closedness kills the finite
  extension to give \(\Gamma(X_y) = \bar k\). A morphism into the affine \(U_0\) is pinned by its ring
  map on global sections (\texttt{ext\_of\_isAffine}), so \(f\) and
  \(\mathtt{retract} \fatsemi f\) agree on each closed slice. Closed points are dense in the
  locally-of-finite-type \(U\) (\texttt{closure\_closedPoints} / the Jacobson-space property of
  locally-of-finite-type \(\bar k\)-schemes; in the formalization this density is supplied by the new
  hypothesis \([\mathtt{LocallyOfFiniteType}\ (X \otimes Y).\mathrm{hom}]\)), so the per-slice
  agreement globalises to all of \(U\) via the reduced-source / separated-target rigidity
  \texttt{ext\_of\_isDominant\_of\_isSeparated'} (the same handle \texttt{rigidity\_core} already
  uses). The packaged ``morphisms agreeing on a dense set of closed points are equal''
  hom-extensionality --- the one connective Mathlib does not supply directly (it gives only the
  single-dominant-morphism \texttt{ext\_of\_isDominant}) --- has now been built as the axiom-clean
  \cref{lem:morphism_eq_of_eqAt_closedPoints} (Step~2): it assembles the closed points into one
  dominant coproduct probe and applies \texttt{ext\_of\_isDominant}. The whole argument is extracted
  into the dedicated obligation \cref{lem:rigidity_eqOn_saturated_open_to_affine} below, whose body
  wires that proven Step~2 over the per-slice Step~1
  (\cref{lem:rigidity_eqAt_closedPoint_of_proper_into_affine}); Step~1 (Mumford's per-closed-slice
  constancy) is now the chain's \emph{single} genuinely-deep residual \texttt{sorry}.
\end{proof}
\begin{lemma}
\leanok
[Bridge~2: slice-constancy on a saturated open mapping into an affine]
  \label{lem:rigidity_eqOn_saturated_open_to_affine}

  \lean{AlgebraicGeometry.rigidity_eqOn_saturated_open_to_affine}
  \uses{lem:rigidity_eqAt_closedPoint_of_proper_into_affine, lem:morphism_eq_of_eqAt_closedPoints}
  % NOTE (iter-161): the iter-160 SIGNATURE GAP is RESOLVED. The Lean proof wires Step 2
  % (morphism_eq_of_eqAt_closedPoints, PROVEN axiom-clean) over Step 1
  % (rigidity_eqAt_closedPoint_of_proper_into_affine, the residual sorry). Step 2 needs the closed
  % points of U to be DENSE, i.e. U Jacobson; over k̄ this follows from (X⊗Y) being locally of finite
  % type, so [LocallyOfFiniteType (X⊗Y).hom] is now a stated hypothesis of this lemma (and across the
  % chain). With it, the `JacobsonSpace U` instance is a routine discharge (LocallyOfFiniteType.
  % jacobsonSpace + JacobsonSpace.of_isOpenEmbedding), no longer an as-typed gap. The only `sorry`
  % reachable from this lemma is now Step 1 (Mumford's per-closed-slice constancy). The hypothesis
  % set is recorded in the statement and proof prose below (both [IsAlgClosed k̄] AND
  % [LocallyOfFiniteType (X⊗Y).hom]); the earlier "[IsAlgClosed] is the only added instance" claim
  % is retired.
  % SOURCE: [Mumford, Abelian Varieties], Rigidity Lemma (Form I.), Ch.~II \S4, book p.~43 (read from references/mumford-abelian-varieties.pdf, PDF page 54)
  % SOURCE QUOTE PROOF: "For each \(y \in V\), the complete variety \(X \times \{y\}\) gets mapped by
  % \(f\) into the affine variety \(U\), and hence to a single point of \(U\). But this means that for
  % any \(x \in X\), \(y \in V\), \(f(x, y) = f(x_0, y) = g \circ p_2(x, y)\), and this proves our
  % assertion." [Mumford, Ch.~II \S4, Rigidity Lemma, p.~43.]
  \textit{Source: Mumford, Abelian Varieties, Ch.~II \S4, Rigidity Lemma (Form~I), p.~43 (the
  ``for each \(y \in V\), the complete slice maps into the affine, hence to a single point'' step).}
  Let \(\bar k\) be algebraically closed and let \(X\), \(Y\), \(Z\) be objects of
  \(\mathrm{Over}\,(\Spec \bar k)\) with \(X\) proper, \(X \times Y\) geometrically irreducible, reduced,
  and \emph{locally of finite type} over \(\bar k\) (the formalization carries
  \([\mathtt{LocallyOfFiniteType}\ (X \otimes Y).\mathrm{hom}]\), which makes the saturated open \(U\)
  below a Jacobson space, so its closed points are dense --- the hypothesis Step~2 of the proof
  consumes), and \(Z\) separated. Let \(f \colon X \times Y \to Z\) be a morphism, fix a \(\bar k\)-point
  \(x_0 \colon \mathbf 1 \to X\), and write \(\mathtt{retract} := \mathtt{lift}\,
  (\mathtt{toUnit}\,(X \times Y) \fatsemi x_0)\,(p_2)\) for the ``\((x, y) \mapsto (x_0, y)\)''
  endomorphism. Suppose:
  \begin{itemize}
    \item \(U \subseteq X \times Y\) is a \(p_2\)-\emph{saturated} open, i.e.\ \(U = p_2^{-1}(V_{\mathrm{set}})\)
      for a set \(V_{\mathrm{set}} \subseteq Y\) (so \(U\) contains the whole fibre \(X_y\) over each of
      its points);
    \item \(U_0 \subseteq Z\) is an \emph{affine} open;
    \item \(f(U) \subseteq U_0\), i.e.\ \(f\) maps \(U\) into the affine \(U_0\).
  \end{itemize}
  Then \(f\) and \(\mathtt{retract} \fatsemi f\) agree on \(U\) as scheme morphisms:
  \(U.\iota \fatsemi f = U.\iota \fatsemi (\mathtt{retract} \fatsemi f)\).
\end{lemma}
\begin{proof}
  \leanok
  This is the formalization realization of Mumford's step ``for each \(y \in V\), the complete
  variety \(X \times \{y\}\) gets mapped by \(f\) into the affine variety \(U\), and hence to a single
  point of \(U\)''. We realise it by a \textbf{global-sections + per-closed-point} route that is
  \emph{cohomology-free}; in particular the relative Stein-factorisation / proper-pushforward
  \(f_*\mathcal O = \mathcal O\) framing is a confirmed Mathlib gap (no Stein factorisation, no
  \(H^0\) base change) and is \emph{deliberately avoided}. The argument splits into two named
  sub-lemmas, one residual and one proven.
  \medskip
  \noindent\textbf{Step 1 (per-closed-slice constancy) --- \cref{lem:rigidity_eqAt_closedPoint_of_proper_into_affine}, the residual.}
  Fix a closed point \(x\) of \(U\), lying over a closed point \(y = p_2(x) \in V_{\mathrm{set}}\).
  Since \(\bar k\) is algebraically closed and \(Y\) is of finite type over \(\bar k\), the residue field
  is \(\kappa(y) = \bar k\). Because \(U = p_2^{-1}(V_{\mathrm{set}})\) is saturated, \(U\) contains the
  \emph{entire} fibre \(X_y \cong X\) over \(y\); hence \(f\) maps the proper integral slice \(X_y\) into the
  affine \(U_0\). By \(\mathtt{isField\_of\_universallyClosed}\) the global sections \(\Gamma(X_y)\) form a
  field, module-finite over \(\bar k\) by \(\mathtt{finite\_appTop\_of\_universallyClosed}\), and
  algebraic closedness collapses the finite extension to give \(\Gamma(X_y) = \bar k\). A morphism into
  the affine \(U_0\) is pinned by its ring map on global sections (\(\mathtt{ext\_of\_isAffine}\)), so the
  restriction of \(f\) to \(X_y\) factors through a single \(\bar k\)-point of \(U_0\) --- necessarily the
  point \(f(x_0, y) = (\mathtt{retract} \fatsemi f)(x)\). Thus \(f\) and \(\mathtt{retract} \fatsemi f\)
  agree (after the residue-field probe) at every closed point of \(U\). This per-slice statement is the
  chain's single genuinely-deep residual \texttt{sorry}.
  \medskip
  \noindent\textbf{Step 2 (globalisation over dense closed points) --- \cref{lem:morphism_eq_of_eqAt_closedPoints}, proven.}
  The closed points are dense in the locally-of-finite-type \(\bar k\)-scheme \(U\)
  (\(\mathtt{closure\_closedPoints}\) / the Jacobson-space property, supplied here by the new hypothesis
  \([\mathtt{LocallyOfFiniteType}\ (X \otimes Y).\mathrm{hom}]\)). To turn ``\(f\) and
  \(\mathtt{retract} \fatsemi f\) agree at each closed point'' into a single morphism equality on \(U\),
  the proven globalisation \cref{lem:morphism_eq_of_eqAt_closedPoints} assembles the closed points
  into one dominant probe --- the coproduct
  \(\coprod_{x \in \mathrm{closedPoints}\,U} \Spec \kappa(x) \to U\), whose range is dense --- and feeds
  it to the reduced-source / separated-target rigidity handle \(\mathtt{ext\_of\_isDominant}\) (the same
  family \(\mathtt{rigidity\_core}\) uses). This yields the morphism equality
  \(U.\iota \fatsemi f = U.\iota \fatsemi (\mathtt{retract} \fatsemi f)\) on all of \(U\). The
  ``agrees on a dense set of closed points \(\Rightarrow\) hom-equality'' connective is the one piece
  Mathlib does not package directly (it supplies only the single-dominant-morphism
  \(\mathtt{ext\_of\_isDominant}\)); that bespoke build is now done in
  \cref{lem:morphism_eq_of_eqAt_closedPoints}, and it is \emph{not} cohomology. \qedhere
\end{proof}
\begin{lemma}\leanok
[Step~2: morphisms agreeing at every closed point of a Jacobson source are equal]
  \label{lem:morphism_eq_of_eqAt_closedPoints}
  
  \lean{AlgebraicGeometry.morphism_eq_of_eqAt_closedPoints}
  % Project-bespoke assembly: NO external source. Mathlib supplies only the single-dominant-morphism
  % ext_of_isDominant; the "dense closed points => hom-extensionality" connective is built here.
  Let \(W\), \(Z\) be schemes with \(W\) \emph{reduced}, \(W\) a \emph{Jacobson space}, and \(Z\)
  \emph{separated}. Let \(g_1, g_2 \colon W \to Z\) be morphisms. If, for every \emph{closed} point
  \(x \in W\), the two morphisms agree after restriction along the residue-field point
  \(\Spec \kappa(x) \to W\) (in Lean, \(W.\mathtt{fromSpecResidueField}\,x \fatsemi g_1 =
  W.\mathtt{fromSpecResidueField}\,x \fatsemi g_2\)), then \(g_1 = g_2\).
\end{lemma}
\begin{proof}
\leanok
  This is the ``dense closed points \(\Rightarrow\) hom-extensionality'' connective that Mathlib does
  not package directly: it supplies only the single-dominant-morphism extensionality
  \(\mathtt{ext\_of\_isDominant}\) (a dominant morphism into a separated target is an epimorphism for
  the purpose of right-cancellation). We assemble the per-closed-point hypotheses into one such
  dominant probe.
  Form the coproduct \(P := \coprod_{x \in \mathrm{closedPoints}\,W} \Spec \kappa(x)\) and the morphism
  \(\mathrm{probe} \colon P \to W\) defined by \(\mathtt{Sigma.desc}\) of the residue-field points
  \(W.\mathtt{fromSpecResidueField}\,x\). Its topological range contains every closed point of \(W\):
  the inclusion \(\iota_x \colon \Spec \kappa(x) \to P\) composes with \(\mathrm{probe}\) to
  \(W.\mathtt{fromSpecResidueField}\,x\) (the defining property of \(\mathtt{Sigma.desc}\)), and the
  range of \(W.\mathtt{fromSpecResidueField}\,x\) is the singleton \(\{x\}\)
  (\(\mathtt{Scheme.range\_fromSpecResidueField}\)). Because \(W\) is a Jacobson space, the closed points
  are dense (\(\mathtt{closure\_closedPoints}\), i.e.\ \(\overline{\mathrm{closedPoints}\,W} = W\)), so
  the range of \(\mathrm{probe}\) is dense and \(\mathrm{probe}\) is \emph{dominant}.
  Componentwise, \(g_1\) and \(g_2\) agree after \(\mathrm{probe}\): precomposing with each coproduct
  inclusion \(\iota_x\) reduces (by \(\mathtt{Sigma.\iota\_desc}\)) to
  \(W.\mathtt{fromSpecResidueField}\,x \fatsemi g_1 = W.\mathtt{fromSpecResidueField}\,x \fatsemi g_2\),
  which is the hypothesis at the closed point \(x\). By the universal property of the coproduct
  (\(\mathtt{Sigma.hom\_ext}\)) this gives \(\mathrm{probe} \fatsemi g_1 = \mathrm{probe} \fatsemi g_2\).
  Since \(\mathrm{probe}\) is dominant and \(Z\) is separated, \(\mathtt{ext\_of\_isDominant}\) cancels it
  on the left to yield \(g_1 = g_2\).
\end{proof}
\begin{lemma}
\leanok
[Global-sections constancy: two \(\bar k\)-points of a proper integral source into an affine agree]
  \label{lem:eq_comp_of_isAffine_of_properIntegral}
  \lean{AlgebraicGeometry.eq_comp_of_isAffine_of_properIntegral}
  % Project-bespoke assembly of Mathlib lemmas: NO external source. Realises, cohomology-free, the
  % classical "a global regular function on a proper integral variety over an algebraically closed
  % field is constant" --- precisely the affine-image step of Mumford's Rigidity Lemma proof quoted
  % above (the "complete slice into an affine maps to a single point" sentence, Ch.~II \S4, p.~43).
  Let \(\bar k\) be algebraically closed and let \(W\) be an integral scheme equipped with a structure
  morphism \(w \colon W \to \Spec \bar k\) that is universally closed (proper) and locally of finite
  type. Let \(g \colon W \to V\) be a morphism into an \emph{affine} scheme \(V\). Then any two
  \(\bar k\)-points of \(W\) --- i.e.\ any two sections \(a, b \colon \Spec \bar k \to W\) of \(w\) (so
  \(a \fatsemi w = \mathbf 1\) and \(b \fatsemi w = \mathbf 1\)) --- satisfy \(a \fatsemi g = b \fatsemi g\).
  \medskip
  \noindent\textbf{Every hypothesis is load-bearing} (confirmed by the iter-161 audit).
  \([\mathtt{IsAlgClosed}\ \bar k]\) collapses the finite extension \(\Gamma(W)/\bar k\); integrality of
  \(W\) makes \(\Gamma(W)\) a domain (without it a two-point disjoint union is a counterexample);
  \(\mathtt{UniversallyClosed}\) is what upgrades \(\Gamma(W)\) to a \emph{field} (without it
  \(\mathbb A^1\), with \(\Gamma = \bar k[t]\) and two distinct \(\bar k\)-points, is a counterexample);
  \(\mathtt{LocallyOfFiniteType}\) gives finiteness of \(\Gamma(W)\) over \(\bar k\); and
  \([\mathtt{IsAffine}\ V]\) is what lets \(\mathtt{ext\_of\_isAffine}\) pin the morphism by its ring map
  on global sections.
\end{lemma}
\begin{proof}
\leanok
  Because \(W\) is integral and \(w\) is universally closed, the global sections
  \(\Gamma(W, \mathcal O_W)\) form a field (\(\mathtt{isField\_of\_universallyClosed}\)), and the structure
  ring map \(w^\sharp \colon \bar k \to \Gamma(W)\) (the map \(\mathtt{wk.appTop}\) after the canonical
  identification \(\Gamma(\Spec \bar k) \cong \bar k\)) is integral / module-finite
  (\(\mathtt{finite\_appTop\_of\_universallyClosed}\)). Since \(\bar k\) is algebraically closed and
  \(\Gamma(W)\) is a domain finite over it, that map is bijective
  (\(\mathtt{IsAlgClosed.ringHom\_bijective\_of\_isIntegral}\)); thus \(\Gamma(W) = \bar k\) and
  \(w^\sharp\) is an isomorphism on global sections --- in particular an epimorphism. Both sections
  \(a\), \(b\) left-invert \(w^\sharp\) on global sections (they are sections of \(w\)), so they induce the
  \emph{same} ring map \(a^\sharp = b^\sharp\) on \(\Gamma(W)\) (cancel the epimorphism \(w^\sharp\)). A
  morphism into the \emph{affine} \(V\) is pinned by its global-sections ring map
  (\(\mathtt{ext\_of\_isAffine}\)), whence \(a \fatsemi g = b \fatsemi g\). This realises, without
  cohomology, the classical ``proper connected variety into an affine is a single point'' step
  quoted from Mumford above.
\end{proof}
\begin{lemma}
\leanok
[Integrality descends to a retract of an integral scheme]
  \label{lem:isIntegral_of_retract_of_integral}
  \lean{AlgebraicGeometry.isIntegral_of_retract}
  % NOTE: (iter-162 review) project-bespoke (elementary topology + algebra); NO external source.
  % LANDED axiom-clean as `AlgebraicGeometry.isIntegral_of_retract`, stated generically for any
  % retract `r : S ⟶ T`, `pr : T ⟶ S` with `r ≫ pr = 𝟙 S` of an integral `T` (more general than
  % the `X`/`X×Y`/`p_1` instance below). PROOF DIVERGENCE (cosmetic, sound either way): the Lean
  % proves reducedness PER-STALK (each `S.stalk x` embeds via `pr.stalkMap` into the reduced
  % `T.stalk`, then `isReduced_of_isReduced_stalk`), not via the global-sections split-injection
  % the prose below describes. Plan/blueprint-writer may align the prose to the stalk route.
  Let \(X \times Y\) be an integral scheme (here it is geometrically irreducible and reduced over
  \(\bar k\)). Suppose \(X\) is a \emph{retract} of \(X \times Y\): there is a section
  \(s \colon X \to X \times Y\) of the first projection \(p_1\), i.e.\ \(s \fatsemi p_1 = \mathbf 1_X\).
  Then \(X\) is integral.
\end{lemma}
\begin{proof}
\leanok
  Integrality is reducedness together with irreducibility, and the retract supplies each.
  \emph{Reduced.} The section identity \(s \fatsemi p_1 = \mathbf 1_X\) makes the pullback on global
  sections \(p_1^\sharp \colon \mathcal O_X(X) \to \mathcal O_{X \times Y}(X \times Y)\) split
  injective (it is left-inverted by \(s^\sharp\)). A split-injective ring map into the reduced ring
  \(\mathcal O_{X \times Y}(X \times Y)\) embeds \(\mathcal O_X(X)\) into a reduced ring, so it too is
  reduced (\(\mathtt{isReduced\_of\_injective}\) on the split injection); hence \(X\) is reduced.
  \emph{Irreducible.} The first projection \(p_1 \colon X \times Y \to X\) is surjective because it
  admits the section \(s\) (every point of \(X\) is \(p_1(s(\cdot))\)). The continuous image of an
  irreducible space under a surjection is irreducible; as \(X \times Y\) is irreducible and \(p_1\) is a
  continuous surjection onto \(X\), the space of \(X\) is irreducible.
  Reduced and irreducible together give that \(X\) is integral. This is elementary --- no cohomology,
  no external citation --- and feeds the Step-1 geometric assembly below, where the proper slice
  \(X_y \cong X\) must be presented as proper \emph{integral}.
\end{proof}
\begin{lemma}\leanok
[Step~1: per-closed-slice constancy of a proper slice into an affine]
  \label{lem:rigidity_eqAt_closedPoint_of_proper_into_affine}

  \lean{AlgebraicGeometry.rigidity_eqAt_closedPoint_of_proper_into_affine}
  \uses{lem:eq_comp_of_isAffine_of_properIntegral, lem:isIntegral_of_retract_of_integral}
  % SOURCE: [Mumford, Abelian Varieties], Rigidity Lemma (Form I.), Ch.~II \S4, book p.~43 (read from references/mumford-abelian-varieties.pdf, PDF page 54)
  % SOURCE QUOTE PROOF: "For each \(y \in V\), the complete variety \(X \times \{y\}\) gets mapped by
  % \(f\) into the affine variety \(U\), and hence to a single point of \(U\). But this means that for
  % any \(x \in X\), \(y \in V\), \(f(x, y) = f(x_0, y) = g \circ p_2(x, y)\), and this proves our
  % assertion." [Mumford, Ch.~II \S4, Rigidity Lemma, p.~43.]
  \textit{Source: Mumford, Abelian Varieties, Ch.~II \S4, Rigidity Lemma (Form~I), p.~43 (the
  ``for each \(y \in V\), the complete slice maps into the affine, hence to a single point'' step).}
  With the data of \cref{lem:rigidity_eqOn_saturated_open_to_affine} --- \(\bar k\) algebraically
  closed, \(X\) proper over \(\bar k\), \(X \times Y\) geometrically irreducible, reduced, and locally of
  finite type, \(Z\) separated, a \(\bar k\)-point \(x_0 \colon \mathbf 1 \to X\), a morphism
  \(f \colon X \times Y \to Z\), a \(p_2\)-saturated open \(U = p_2^{-1}(V_{\mathrm{set}})\) mapping into an
  affine open \(U_0 \subseteq Z\) --- fix a \emph{closed} point \(x\) of \(U\). Then \(f\) and the collapsed
  map \(\mathtt{retract} \fatsemi f\) (with \(\mathtt{retract} := \mathtt{lift}\,
  (\mathtt{toUnit}\,(X \times Y) \fatsemi x_0)\,(p_2)\), i.e.\ \((x, y) \mapsto (x_0, y)\)) agree at \(x\)
  after the residue-field probe \(\Spec \kappa(x) \to U\):
  \[ U.\mathtt{fromSpecResidueField}\,x \fatsemi (U.\iota \fatsemi f)
     = U.\mathtt{fromSpecResidueField}\,x \fatsemi (U.\iota \fatsemi (\mathtt{retract} \fatsemi f)). \]
\end{lemma}
\begin{proof}
  \leanok
  This is Mumford's per-slice step, realised \emph{cohomology-free} (route~B): the relative Stein /
  proper-pushforward \(f_*\mathcal O = \mathcal O\) framing is a confirmed Mathlib gap and is
  deliberately avoided. The closed point \(x\) of \(U\) lies over a closed point
  \(y = p_2(x) \in V_{\mathrm{set}}\); since \(\bar k\) is algebraically closed and \(Y\) is of finite type
  over \(\bar k\), the residue field is \(\kappa(y) = \bar k\). Saturation \(U = p_2^{-1}(V_{\mathrm{set}})\)
  puts the \emph{entire} proper integral fibre \(X_y \cong X\) over \(y\) inside \(U\), so \(f\) maps \(X_y\)
  into the affine \(U_0\). Because \(X_y\) is proper integral over \(\bar k\), its global sections form a
  field (\(\mathtt{isField\_of\_universallyClosed}\)), module-finite over \(\bar k\)
  (\(\mathtt{finite\_appTop\_of\_universallyClosed}\)); algebraic closedness collapses the finite
  extension, giving \(\Gamma(X_y, \mathcal O) = \bar k\).
  % NOTE (iter-162): the iter-161 documentary gap is RESOLVED --- this algebraic core ("two k-bar
  % points of a proper integral l.f.t. k-bar-scheme into an affine agree") now has its own node
  % lem:eq_comp_of_isAffine_of_properIntegral (above), and this proof's  edge points at it.
  A morphism into the affine \(U_0\) is pinned by
  its ring map on global sections (\(\mathtt{ext\_of\_isAffine}\)), so \(f|_{X_y}\) factors through a
  single \(\bar k\)-point of \(U_0\). That point is \(f(x_0, y)\), since the \(\bar k\)-point \((x_0, y)\) lies
  on \(X_y\); and \(f(x_0, y) = (\mathtt{retract} \fatsemi f)(x)\) by the definition of \(\mathtt{retract}\).
  Evaluating at the residue-field probe of the closed point \(x\) gives the asserted equality. This
  per-slice statement is the chain's single genuinely-deep residual \texttt{sorry}; its globalisation
  over the dense closed points is \cref{lem:morphism_eq_of_eqAt_closedPoints}.
\end{proof}
\section{The Milne \S I.1--I.3 corollaries: additivity, homomorphisms, extension}
This section assembles, from the proven Rigidity Lemma, the structural corollaries used by the
Albanese construction. The load-bearing ingredient is the additive decomposition over a product
(\cref{lem:hom_additivity_over_product}, Milne Corollary~1.5): a morphism \(h \colon V \times W
\to A\) with \(h(v_0, w_0) = 0\) decomposes uniquely as \(h = f \fatsemi p + g \fatsemi q\). From
it follows that a pointed regular map of abelian varieties is a homomorphism
(\cref{lem:av_regular_map_is_hom}, Milne Corollary~1.2) --- the corollary the positive-genus
Albanese universal property consumes. None of this uses the theorem of the cube. For Route~A,
Milne's Theorem~3.2 (rational-map extension, \cref{thm:rational_map_to_av_extends}) is retained
in this chapter as the remaining Albanese input.
\begin{remark}
  \label{rmk:cube_not_needed}
  \textbf{The theorem of the cube is not needed for these corollaries.} The Milne \S I.1--I.3
  corollaries used here --- additivity over a product (Corollary~1.5,
  \cref{lem:hom_additivity_over_product}) and ``a pointed regular map of abelian varieties is a
  homomorphism'' (Corollary~1.2, \cref{lem:av_regular_map_is_hom}) --- follow directly from the
  Rigidity Theorem~1.1 (our \cref{thm:rigidity_lemma}), with no appeal to the theorem of the
  cube. The theorem of the cube (Milne \S I.5) enters Milne's theory only later, for
  projectivity, the seesaw/square principle, dual abelian varieties, polarizations, and the
  Poincar\'e sheaf. For Route~A: the Albanese universal property that the positive-genus arm
  needs (Milne Proposition~6.1/6.4, \S III.6) is \emph{also} cube-free, resting on Theorem~3.2
  (\cref{thm:rational_map_to_av_extends}) together with Corollaries~1.2 and~1.5 --- which is
  precisely why Theorem~3.2 and Corollary~1.2 are retained in this chapter as Route-A inputs.
\end{remark}
\begin{lemma}
\leanok
[Additive decomposition over a product (Milne Corollary 1.5)]
  \label{lem:hom_additivity_over_product}
  
  \uses{thm:rigidity_lemma}
  \lean{AlgebraicGeometry.hom_additive_decomp_of_rigidity}
  % NOTE: (iter-163 review) The Lean signature carries three product-level instance
  % hypotheses on V⊗W not spelled out in the prose: [GeometricallyIrreducible (V⊗W).hom],
  % [LocallyOfFiniteType (V⊗W).hom], [IsReduced (V⊗W).left] — i.e. "V×W is a variety".
  % These are mathematically free for genuine complete varieties V, W, but Mathlib does not
  % auto-derive the product instances from the factors, so they are carried explicitly to keep
  % the lemma axiom-clean. Also: the Lean states the EXISTENCE half only (canonical
  % axis-restriction witnesses inlined; uniqueness not formalized) and requires only V (not W)
  % complete — strictly more general than the prose. (lean-vs-blueprint-checker avr-iter163.)
  % NOTE: (iter-164 review) iter-164 hygiene dropped the A-side instances [Smooth A.hom] and
  % [GeometricallyIrreducible A.hom] from the Lean signature; the abelian-variety target now
  % requires only [GrpObj A] [IsProper A.hom]. The Lean signature is therefore STRICTLY MORE
  % GENERAL than the prose ("Let A be an abelian variety"): every situation the prose covers is
  % covered, and additionally any proper group object suffices. Axiom-clean preserved.
  % (lean-vs-blueprint-checker avr-iter164.)
  % SOURCE: [Milne, Abelian Varieties], Corollary 1.5, \S I.1, doc p.~9 (read from references/abelian-varieties.pdf, PDF page 15)
  % SOURCE QUOTE: "COROLLARY 1.5. Let V and W be complete varieties over k with k-rational
  % points v0 and w0, and let p and q be the projection maps V×W→V and V×W→W. Let A be an
  % abelian variety. Then a morphism h:V×W→A such that h(v0,w0)=0 can be written uniquely as
  % h = f∘p + g∘q with f:V→A and g:W→A morphisms such that f(v0)=0 and g(w0)=0."
  \textit{Source: Milne, Abelian Varieties, Corollary~1.5.}
  Let \(V\), \(W\) be complete varieties over \(\bar k\) with \(\bar k\)-points \(v_0 \in V\),
  \(w_0 \in W\), and let \(p \colon V \times W \to V\), \(q \colon V \times W \to W\) be the
  projections. Let \(A\) be an abelian variety over \(\bar k\). Then a morphism
  \(h \colon V \times W \to A\) with \(h(v_0, w_0) = 0\) can be written uniquely as
  \(h = f \circ p + g \circ q\), with morphisms \(f \colon V \to A\), \(g \colon W \to A\) satisfying
  \(f(v_0) = 0\) and \(g(w_0) = 0\).
  \emph{Lean encoding.} In the project \(A\) is a group object \texttt{[GrpObj A]} (not assumed
  commutative). The ``\(+\)'' here is the binary operation that \texttt{GrpObj} induces on each
  hom-set \(\mathrm{Hom}(X, A)\): for \(u, v \colon X \to A\), \(u + v\) is the composite
  \(X \xrightarrow{\langle u, v\rangle} A \times A \xrightarrow{\;\mathrm{mul}\;} A\) (and
  \(f \circ p + g \circ q\) is built this way from the projections). The decomposition is therefore
  read \emph{multiplicatively} as \(h = (f \circ p) \cdot (g \circ q)\) in the group
  \(\mathrm{Hom}(V \times W, A)\), and does \emph{not} require \(A\) commutative; commutativity of
  abelian varieties (Milne Cor 1.4, a separate consequence of \cref{lem:av_regular_map_is_hom})
  is needed only to rewrite the product symmetrically/additively, not for the statement or its
  proof. The prover should infer the hom-set group operation from the \texttt{GrpObj} instance,
  not from a \texttt{CommGrpObj} instance.
\end{lemma}
% SOURCE QUOTE PROOF: "PROOF. Set f=h|V×{w0}, g=h|{v0}×W, and identify V×{w0} and {v0}×W with
% V and W. On points, f(v) = h(v,w0) and g(w) = h(v0,w), and so φ =df h − (f∘p + g∘q) is the
% map that sends (v,w) ↦ h(v,w) − h(v,w0) − h(v0,w). Thus φ(V×{w0}) = 0 = φ({v0}×W) and so the
% theorem shows that φ = 0." [Milne, Cor 1.5, \S I.1, doc p.~9--10.]
\begin{proof}
  \leanok
  \textit{Source: Milne, Abelian Varieties, Corollary~1.5.}
  Define \(f := h|_{V \times \{w_0\}}\) and \(g := h|_{\{v_0\} \times W}\), identifying
  \(V \times \{w_0\} \cong V\) and \(\{v_0\} \times W \cong W\); on points \(f(v) = h(v, w_0)\) and
  \(g(w) = h(v_0, w)\). Form the difference
  \[ \varphi := h - (f \circ p + g \circ q) \colon V \times W \to A, \qquad
     \varphi(v, w) = h(v, w) - h(v, w_0) - h(v_0, w) \]
  (the subtraction is in the group \(A\), i.e.\ \(\varphi\) is the composite of
  \((h, -(f \circ p + g \circ q))\) with the group law of \(A\)). Then \(\varphi\) vanishes on both
  axes: \(\varphi(V \times \{w_0\}) = 0 = \varphi(\{v_0\} \times W)\) (direct computation, using
  \(h(v_0, w_0) = 0\)). In particular \(\varphi\) collapses the complete axis \(V \times \{w_0\}\) to
  the single point \(0 \in A\), so by the Rigidity Lemma (\cref{thm:rigidity_lemma}, with \(V\)
  complete and \(V \times W\) geometrically irreducible) \(\varphi\) factors through the projection
  \(q\); but it also vanishes on \(\{v_0\} \times W\), which is a section of \(q\), so the factoring
  morphism is identically \(0\), i.e.\ \(\varphi \equiv 0\). Hence \(h = f \circ p + g \circ q\).
  Uniqueness: restricting any such decomposition to the axes recovers \(f\) and \(g\) as the
  axis-restrictions of \(h\).
\end{proof}
\begin{lemma}
\leanok
[A pointed regular map of abelian varieties is a homomorphism (Milne Corollary 1.2)]
  \label{lem:av_regular_map_is_hom}

  \uses{lem:hom_additivity_over_product}
  \lean{AlgebraicGeometry.av_regularMap_isHom_of_zero}
  % NOTE: (iter-163 review) The Lean formalizes the POINTED direction
  % (η[A] ≫ α = η[B] ⟹ IsMonHom α), not the full "composite of a homomorphism with a
  % translation" form — which is what the genus-0 consumer needs. The conclusion IsMonHom α
  % (one_hom + mul_hom) is a faithful, slightly stronger rendering of "α is a homomorphism".
  % As with lem:hom_additivity_over_product, the signature carries the product-level instances
  % [GeometricallyIrreducible (A⊗A).hom], [LocallyOfFiniteType (A⊗A).hom], [IsReduced (A⊗A).left]
  % ("A×A is a variety"), free for a genuine abelian variety A but carried explicitly because
  % Mathlib lacks the product-instance inference. (lean-vs-blueprint-checker avr-iter163.)
  % NOTE: (iter-164 review) iter-164 hygiene dropped the B-side instances [Smooth B.hom] and
  % [GeometricallyIrreducible B.hom] from the Lean signature; the target B now requires only
  % [GrpObj B] [IsProper B.hom]. (The source-side A retains the full abelian-variety instance
  % set.) The Lean signature is strictly more general on B than the prose ("α : A → B of
  % abelian varieties") demands. The lean-auditor flagged the symmetric A-side
  % [Smooth A.hom]/[GeometricallyIrreducible A.hom] as ALSO unused — kept this iter as a
  % minimal-change follow-up. (lean-vs-blueprint-checker avr-iter164.)
  % SOURCE: [Milne, Abelian Varieties], Corollary 1.2, \S I.1, doc p.~9 (read from references/abelian-varieties.pdf, PDF page 15)
  % SOURCE QUOTE: "COROLLARY 1.2. Every regular map α:A→B of abelian varieties is the composite
  % of a homomorphism with a translation."
  \textit{Source: Milne, Abelian Varieties, Corollary~1.2.}
  Every regular map \(\alpha \colon A \to B\) of abelian varieties over \(\bar k\) is the composite
  of a homomorphism with a translation; equivalently, if \(\alpha(0_A) = 0_B\) then \(\alpha\) is a
  homomorphism, i.e.\ \(\alpha(a + a') = \alpha(a) + \alpha(a')\) for all \(a, a' \in A\).
\end{lemma}
% SOURCE QUOTE PROOF: "PROOF. The regular map α will send the k-rational point 0 of A to a
% k-rational point b of B. After composing α with translation by −b, we may assume that
% α(0)=0. Consider the map φ:A×A→B, φ(a,a′)=α(a+a′) − α(a) − α(a′). By this we mean that φ is
% the difference of the two regular maps A×A →(m)→ A →(α)→ B and A×A →(α×α)→ B×B →(m)→ B, which
% is a regular map. Then φ(A×0)=0=φ(0×A) and so φ=0. This means that α is a homomorphism."
% [Milne, Cor 1.2, \S I.1, doc p.~9.]
\begin{proof}
  \leanok
  \textit{Source: Milne, Abelian Varieties, Corollary~1.2.}
  The image \(\alpha(0_A)\) is a \(\bar k\)-point \(b\) of \(B\); composing \(\alpha\) with translation by
  \(-b\) we may assume \(\alpha(0_A) = 0_B\), and it suffices to show such an \(\alpha\) is a
  homomorphism. Form
  \[ \varphi \colon A \times A \to B, \qquad \varphi(a, a') = \alpha(a + a') - \alpha(a) - \alpha(a'), \]
  the difference of the two regular maps \(A \times A \xrightarrow{m} A \xrightarrow{\alpha} B\)
  and \(A \times A \xrightarrow{\alpha \times \alpha} B \times B \xrightarrow{m} B\). Then
  \(\varphi(0_A, 0_A) = 0\) and \(\varphi\) vanishes on both axes,
  \(\varphi(A \times 0) = 0 = \varphi(0 \times A)\) (since \(\alpha(0) = 0\)). By the additive
  decomposition \cref{lem:hom_additivity_over_product} (with \(V = W = A\)), the axis-restrictions
  of \(\varphi\) being \(0\) force \(\varphi \equiv 0\); equivalently, \(\varphi\) collapses the complete
  axis \(A \times 0\) and vanishes on \(0 \times A\), so Rigidity gives \(\varphi = 0\). Hence
  \(\alpha(a + a') = \alpha(a) + \alpha(a')\), i.e.\ \(\alpha\) is a homomorphism.
\end{proof}
\begin{remark}
  \label{rmk:thm32_codim1_mathlib_gap}
  \textbf{Formalisation risk: the pure-codimension-\(1\) step (Lemma 3.3) has no Mathlib
  Weil-divisor support.} Theorem~3.1's everywhere-extension-outside-codimension-\(2\) is within
  reach of Mathlib's valuative criterion of properness
  (\texttt{ValuativeCriterion}/\texttt{IsProper.of\_valuativeCriterion}), applied at the
  discrete valuation ring \(\mathcal O_{V, Z}\) of each codimension-\(1\) point \(Z\). But Lemma~3.3,
  which rules out codimension-\(1\) indeterminacy for maps into a group variety, is proved in
  Milne via Weil-divisor theory (\(\mathrm{div}(f)_0 - \mathrm{div}(f)_1\), prime divisors,
  tangent cones) for which Mathlib currently has \emph{no} usable API at this generality. This
  is the route's riskiest sub-build. An open question for the prover: whether a \emph{pointwise}
  valuative extension --- showing, at each codimension-\(1\) point \(Z\) separately, that the
  morphism \(\Spec k(V) \to A\) extends over \(\Spec \mathcal O_{V, Z}\) using only that \(A\) is
  proper and a group (so the difference-map argument can be run \emph{locally} at \(Z\) without
  globally building \(\mathrm{div}(f)\)) --- side-steps building divisor theory entirely.
  \textbf{This surface extension is confined to Route~A.} It is needed only for the
  positive-genus Albanese universal property (Milne \S III.6), where the additive-defect map of a
  morphism out of a higher-dimensional variety must be extended over a complete surface.
  The additive decomposition \cref{lem:hom_additivity_over_product} and the rigidity corollaries
  established above do \emph{not} depend on it. Consequently
  \cref{thm:rational_map_to_av_extends} (Theorem~3.2), the codim-\(1\) indeterminacy sub-build
  (Lemma~3.3), and the Weil-divisor gap discussed above are all confined to \textbf{Route~A};
  they are retained in this chapter solely as Route-A inputs and may be deferred until Route~A
  consumes them.
\end{remark}
